---
title: "Race/ethnicity in YRBS"
author: "Thomas Lumley"
date: "03/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `usethnicity` data set contains variables on race and ethnic identification from the 2017 Youth Risk Behaviour Survey, together with two variables on smoking behaviour.  The YRBS is a multistage cluster-sampled survey, so valid inference about associations requires using survey design information. This subset of variables without weights is useful only for demonstration purposes. 

```{r}
library(rimu)
data(usethnicity)
head(usethnicity)
```

Question 4 asks *Are you Hispanic or Latino?*, and Question 5 asks for any of  

- A. American Indian or Alaska Native, 
- B. Asian, 
- C. Black or African American, 
- D. Native Hawaiian or Other Pacific Islander, 
- E. White

that apply.  In the data set, these five letters are pasted together into a single variable.



We need to split `Q5` into its component letters. One way is to use `strsplit`, producing a list with a vector of zero to five letters for each individual, then call the `as.mr` method for lists.

```{r}
race<-as.mr(strsplit(as.character(usethnicity$Q5),""))
mtable(race)
```

There's a spurious `" "` category from the string splitting, and the values `F`, `G`, and `H` are also invalid, so we need to remove them

```{r}
race<-mr_drop(race,c(" ","F","G","H"))
mtable(race)
```

We might want easier-to-recognise names for the categories

```{r}
race <- mr_recode(race, AmIndian="A",Asian="B", Black="C", Pacific="D", White="E")
```

Now, Hispanic/Latino ethnicity is asked in a separate question. We convert it via the `as.mr` method for logical vectors, and then combine it with `race`

```{r}
hispanic<-as.mr(usethnicity$Q4==1, "Hispanic")
ethnicity<-mr_union(race, hispanic)
ethnicity[101:120]
```

The `plot` method shows co-occurence of the various race/ethnicity terms

```{r}
plot(ethnicity,nsets=6)
```

Tabulations against other factor or multiple-response variables are possible with `mtable`.  Note that `mtable` shows frequencies for each category; use `as.character` to get frequencies for combinations -- do not use `as.factor`, which is not generic and so cannot have a `mr` method.

```{r}
mtable(ethnicity, usethnicity$QN30)
table(ethnicity %has% "Black", usethnicity$QN30)
table(ethnicity %hasonly% "Black", usethnicity$QN30)
table(as.character(ethnicity), usethnicity$QN30)
table(forcats::fct_lump(as.character(ethnicity), n=10),    usethnicity$QN30)
```