---
title: "RStudio Community Survey"
author: "Annie Cohen"
date: "9 March 2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RStudio Community Survey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rimu)
```

The `rstudiosurvey` data set contains 1838 rows of responses from the 2019 RStudio Community Survey, where columns are the 51 questions and a column for the timestamp. A few of the questions included in the survey are:

```{r message = FALSE, warning = FALSE}
data("rstudiosurvey")
colnames(rstudiosurvey)[16:22]
```

Perhaps we are interested in understanding which R applications are used by the R community. We might be interested in the response to *"What applications do you use R for most? (check all that appply)"*. Notice that this question allows for multiple responses:

- Statistical analysis
- Data transformation
- Modeling
- Visualization
- Machine learning
- Text processing
- Genomics
- Medicine
- Survey analysis
- Clinical trials
- Financial analysis
- Other

We can convert this column into a multiple response object using `as.td`:

```{r}
apps <- as.td(strsplit(rstudiosurvey$`What applications do you use R for most? (check all that apply)`, ","))

tdtable(apps)[,1:20]
```

Yikes, this is only the first 20. There are 169 total unique entries for this question. In order to make things easier for our analysis, we can keep only entries that aren't created by the `Other` option using `td_drop` to specify the columns we would like to drop:

```{r}
new_apps <- td_drop(apps, colnames(tdtable(apps))[-c(1:20)])
tdtable(new_apps)
```

This is still a lot of . We can use `td_lump` to choose only the values that appear in at least 10% of the responses:

```{r}
library(vctrs)
common_apps <- td_lump(apps, prop=.1)
tdtable(common_apps)
```

That's better. Now we can look at the co-occurrence of the responses with `table()` and the `plot` method:

```{r fig.width = 6, fig.height=4}
tdtable(common_apps, common_apps)
plot(common_apps, nsets = 9)
```

The object `common_apps` is a `vctrs_vctr` which allows it to be integrated into data frames and manipulated with dplyr functions. We can store it back into the data frame `rstudiosurvey` as a new column:

```{r}
tidy_survey <- rstudiosurvey %>%
  mutate(`R applications (tidy)` = common_apps)

print(tidy_survey$`R applications (tidy)`[1:6])

ncol(tidy_survey)
```

Looks good, but that's a lot of columns! Let's use `dplyr` to select some of the important survey questions. Because the multiple response `td` object is situated as a column within a data frame, we can use `tidyverse` functions to manipulate the data frame.

```{r}
smaller <- tidy_survey %>%
  select(Timestamp,
         `How likely are you to recommend R to a colleague, friend, or family member?`,
         `How did you learn R? If you used multiple methods, please select the one you used the most.`,
         `What do you like best about using R?`,
         `How would you rate your level of experience using R?`,
         `R applications (tidy)`)

glimpse(smaller)
```

Now, let's compare R applications with the reported level of experience of the respondent.

```{r}
x <- data.frame(tdtable(smaller$`R applications (tidy)`, smaller$`How would you rate your level of experience using R?`))

round(x/rowSums(x),3) %>%
  select("Beginner","Intermediate","Expert")
```

From this, we can see that there are very few beginners who are doing financial analysis, modeling, text processing, or machine learning. Furthermore, most R users identify as intermediate. We can also see that those who use machine learning applications identify as an expert almost as often as they do an intermediate, which is not the case for other applications. Using `tdtable` to compare two different responses in the RStudio Community Survey, we can gain insight into the relationship between reported experience level and R applications used.

```{r}
tidyr::pivot_longer(smaller$`R applications (tidy)`)
```
