---
title: "RStudio Community Survey"
author: "Annie Cohen"
date: "01/19/2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RStudio Community Survey}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rimu)
library(dplyr)
library(vctrs)
```

The `rstudiosurvey` data set contains 1811 rows of responses from the 2019 RStudio Community Survey, where columns are the 51 questions and a column for the timestamp. A few of the questions included in the survey are

```{r message = FALSE, warning = FALSE}
data("rstudiosurvey")
colnames(rstudiosurvey)[16:22]
```

Perhaps we are interested in understanding which R applications are used by the R community. We might be interested in the response to "What applications do you use R for most? (check all that appply)". Notice that this question allows for multiple responses:

- Statistical analysis
- Data transformation
- Modeling
- Visualization
- Machine learning
- Text processing
- Genomics
- Medicine
- Survival analysis
- Clinical trials
- Financial analysis

So we can convert this column into a multiple response object using `as.td`

```{r}
rimu::as.td(strsplit(rstudiosurvey$`What applications do you use R for most? (check all that apply)`, ","))
```









## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
